Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-28 cvx-flats-neurips21/cvx-flats.tex
final

%\usetikzlibrary{calc}
\newcommand{\Comments}{1}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
	\todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{teal}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{teal!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{}%\mytodo{orange!80!white}{BTW: #1}}
\newcommand{\discuss}[1]{\mynote{cyan!20!white}{#1}}
\ifnum\Comments=1               % fix margins for todonotes
\setlength{\marginparwidth}{1in}
\fi

@@

\paragraph{Literature on consistency and calibration.}% \label{subsec:consistency}
%\bo{Citation help appreciated!} \jessie{\cite{fisher1922mathematical,zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex} -- these are the ones I know of that discuss consistency in slightly various forms.  I believe the definition we run with is most similar to \cite{ramaswamy2016convex}.}
%\jessiet{Make this subsection more concise?}
We focus on surrogate losses $L: \reals^d \times \Y \to \reals$ that are consistent, roughly meaning that minimizing $L$-loss corresponds to solving the target problem of interest.
When given a \emph{target loss} $\ell$, we roughly define $L$ to be consistent if minimizing $L$, and applying a link function, minimizes $\ell$ (Def.~\ref{def:consistent-ell}, \S~\ref{sec:consis-implies-indir})~\citep{zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex}.
When given instead a target statistic such as the conditional quantile or variance, we introduce a notion of consistency in line with classical statistics (Def.~\ref{def:consistent-prop}, \S~\ref{sec:consis-implies-indir}) \citep{gyorfi2006distribution, fan1998efficient,ruppert1997local}.
Here we define $L$ to be consistent if minimizing $L$ and applying a link function yields estimates converging to the correct statistic value.
%\jessiet{Do we still want this following paragraph?}
%\raft{Good question.  I edited a bit, but still up for discussion.}
%A priori, it is not clear that compatible definitions of consistency could be given for both target statistics and target losses.
The key observation which underpins our approach is that consistency for target losses is a special case of consistency for target statistics (\S~\ref{sec:consis-implies-indir}).
Therefore, property elicitation---which studies the exact minimizers of loss functions---allows us to give general lower bounds on prediction dimension of any convex surrogates corresponding to a target task; these bounds apply across all four quadrants.
See \S~\ref{subsec:complexity} for a review of previous work pertaining to notions of prediction dimension.

\begin{table}[h!]
  \begin{center}
    \aboverulesep=1ex
    \belowrulesep=1ex
    \begin{tabular}{p{12ex}ll}
      & \emph{Target loss}  & \emph{Target statistic}\\
      \cmidrule[1pt]{2-3}
%      \parbox{12ex}{\emph{Discrete \\ prediction}} & \textbf{Q1:} Classification  & \textbf{Q2:} Hierarchical classification variant\\ 
      \parbox{12ex}{\emph{Discrete \\ prediction}} & \textbf{Q1}, e.g. classification  & \textbf{Q2}, e.g. hierarchical classification\\ 
      \cmidrule{2-3}
%      \parbox{12ex}{\emph{Continuous \\ estimation}} & \textbf{Q3:} Least-squares regression & \textbf{Q4:} Variance estimation\\ 
      \parbox{12ex}{\emph{Continuous \\ estimation}} & \textbf{Q3}, e.g. least-squares regression & \textbf{Q4}, e.g. variance estimation\\ 
      \cmidrule[1pt]{2-3}
      \\
    \end{tabular}
    \caption{The four quadrants of problem types, with an example for each as discussed in \S~\ref{sec:illustration}.}
    \label{tab:quadrants}
  \end{center}
\end{table}

\begin{figure}[b!]
  \centering
  \begin{tikzpicture}[yscale=0.6,xscale=0.46]

@@

    % edges
    \path[->,bend angle=10]
    (consistent) edge[<->] node[left]{\cite{bartlett2006convexity,steinwart2008support}} (calibrated)
    (consistent) edge[bend left] node[above, sloped]{Prop.~\ref{thm:consistent-implies-indir-elic}} (indir-elic)    
    (calibrated) edge[bend right] (indir-elic)
    (indir-elic) edge node[above, sloped]{Thm.~\ref{thm:convex-flats-inf-dim}} (d-flats)
    (d-flats) edge[bend right] node[below, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds)
    (d-flats) edge[bend left] node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
  \end{tikzpicture}
\caption{Flow and implications of our results. Compared to calibration, we suggest indirect elicitation as a simpler but almost-as-powerful necessary condition for consistency. In particular, we obtain a testable condition (Theorem~\ref{thm:convex-flats-inf-dim}), based on $d$-flats, for the existence of a $d$-dimensional consistent convex surrogate. This condition recovers and strengthens existing calibration-based results for Q1, while simultaneously applying to other quadrants. We illustrate the breadth and power of $d$-flats by resolving two open questions for Q3 and Q4 in \S~\ref{sec:contin-consis}.}
\label{fig:results-flow}
\end{figure}

As definitions of consistency are relatively intractable to apply directly, the literature often focuses on a weaker condition called \emph{calibration}, which only applies when given a target loss, e.g. Quadrants 1 and 3. 
Particularly, several authors~\citep{zhang2004statistical, lin2004note,bartlett2006convexity,tewari2007consistency,ramaswamy2016convex} show the equivalence of consistency and calibration in Quadrant 1, where one is given a target statistic and discrete prediction set.
We discuss the additional relationship of elicitation and calibration in \S~\ref{app:calibration}, and derive Theorem~\ref{thm:consistent-implies-indir-elic} via calibration.

@@

@@

%\begin{ack}
%adam, nishant, nsf grants/fellowship
%\end{ack}

>>>>

%\usetikzlibrary{calc}
\newcommand{\Comments}{0}
\newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
\newcommand{\mytodo}[2]{\ifnum\Comments=1%
	\todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
\newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
\newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
\newcommand{\jessie}[1]{\mynote{teal}{[JF: #1]}}
\newcommand{\jessiet}[1]{\mytodo{teal!20!white}{JF: #1}}
\newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
\newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
\newcommand{\btw}[1]{}%\mytodo{orange!80!white}{BTW: #1}}
\newcommand{\discuss}[1]{\mynote{cyan!20!white}{#1}}
\ifnum\Comments=1               % fix margins for todonotes
\setlength{\marginparwidth}{1in}
\fi

@@

\paragraph{Literature on consistency and calibration.}% \label{subsec:consistency}
%\bo{Citation help appreciated!} \jessie{\cite{fisher1922mathematical,zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex} -- these are the ones I know of that discuss consistency in slightly various forms.  I believe the definition we run with is most similar to \cite{ramaswamy2016convex}.}
%\jessiet{Make this subsection more concise?}
We focus on surrogate losses $L: \reals^d \times \Y \to \reals$ that are consistent, roughly meaning that minimizing $L$-loss corresponds to solving the target problem of interest.
When given a \emph{target loss} $\ell$, we roughly define $L$ to be consistent if minimizing $L$, and applying a link function, minimizes $\ell$ (Def.~\ref{def:consistent-ell}, \S~\ref{sec:consis-implies-indir})~\citep{zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex}.
When given instead a target statistic such as the conditional quantile or variance, we introduce a notion of consistency in line with classical statistics (Def.~\ref{def:consistent-prop}, \S~\ref{sec:consis-implies-indir}) \citep{gyorfi2006distribution, fan1998efficient,ruppert1997local}.
Here we define $L$ to be consistent if minimizing $L$ and applying a link function yields estimates converging to the correct statistic value.
%\jessiet{Do we still want this following paragraph?}
%\raft{Good question.  I edited a bit, but still up for discussion.}
%A priori, it is not clear that compatible definitions of consistency could be given for both target statistics and target losses.
The key observation which underpins our approach is that consistency for target losses is a special case of consistency for target statistics (\S~\ref{sec:consis-implies-indir}).
\begin{table}[h!]
  \vspace*{10pt}
  \begin{center}
    \aboverulesep=1ex
    \belowrulesep=1ex
    \begin{tabular}{p{12ex}ll}
      & \emph{Target loss}  & \emph{Target statistic}\\
      \cmidrule[1pt]{2-3}
%      \parbox{12ex}{\emph{Discrete \\ prediction}} & \textbf{Q1:} Classification  & \textbf{Q2:} Hierarchical classification variant\\ 
      \parbox{12ex}{\emph{Discrete \\ prediction}} & \textbf{Q1}, e.g. classification  & \textbf{Q2}, e.g. hierarchical classification\\ 
      \cmidrule{2-3}
%      \parbox{12ex}{\emph{Continuous \\ estimation}} & \textbf{Q3:} Least-squares regression & \textbf{Q4:} Variance estimation\\ 
      \parbox{12ex}{\emph{Continuous \\ estimation}} & \textbf{Q3}, e.g. least-squares regression & \textbf{Q4}, e.g. variance estimation\\ 
      \cmidrule[1pt]{2-3}
      \\
    \end{tabular}\vspace*{-5pt}
    \caption{The four quadrants of problem types, with an example for each as discussed in \S~\ref{sec:illustration}.\vspace*{-10pt}}
    \label{tab:quadrants}
  \end{center}
\end{table}
\begin{figure}[h!]
  \centering
  \begin{tikzpicture}[yscale=0.6,xscale=0.46]

@@

    % edges
    \path[->,bend angle=10]
    (consistent) edge[<->] node[left]{\cite{bartlett2006convexity,steinwart2008support}} (calibrated)
    (consistent) edge[bend left] node[above, sloped]{Prop.~\ref{thm:consistent-implies-indir-elic}} (indir-elic)    
    (calibrated) edge[bend right] (indir-elic)
    (indir-elic) edge node[above, sloped]{Thm.~\ref{thm:convex-flats-inf-dim}} (d-flats)
    (d-flats) edge[bend right] node[below, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds)
    (d-flats) edge[bend left] node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
  \end{tikzpicture}
\caption{Flow and implications of our results. Compared to calibration, we suggest indirect elicitation as a simpler but almost-as-powerful necessary condition for consistency. In particular, we obtain a testable condition (Theorem~\ref{thm:convex-flats-inf-dim}), based on $d$-flats, for the existence of a $d$-dimensional consistent convex surrogate. This condition recovers and strengthens existing calibration-based results for Q1, while simultaneously applying to other quadrants. We illustrate the breadth and power of $d$-flats by resolving two open questions for Q3 and Q4 in \S~\ref{sec:contin-consis}.}
\label{fig:results-flow}
\end{figure}
Therefore, property elicitation---which studies the exact minimizers of loss functions---allows us to give general lower bounds on prediction dimension of any convex surrogates corresponding to a target task; these bounds apply across all four quadrants.
paragraph
See \S~\ref{subsec:complexity} for a review of previous work pertaining to notions of prediction dimension.

As definitions of consistency are difficult to apply directly, the literature often focuses on a weaker condition called \emph{calibration}, which only applies when given a target loss, e.g. Quadrants 1 and 3. 
Particularly, several authors~\citep{zhang2004statistical, lin2004note,bartlett2006convexity,tewari2007consistency,ramaswamy2016convex} show the equivalence of consistency and calibration in Quadrant 1.
%, where one is given a target statistic \raf{loss?} and discrete prediction set
We discuss the additional relationship of elicitation and calibration in \S~\ref{app:calibration}, and derive Theorem~\ref{thm:consistent-implies-indir-elic} via calibration.

@@

\bigskip
\paragraph{Broader impacts via influence on downstream research.}
\jessiet{Taken from original BI statement for this paper}
This paper takes a step toward an impactful general program: to understand and design \emph{useful} (i.e. convex, consistent, calibrated, etc.) surrogate losses of low dimensionality.
The theory in this paper may help practitioners in a number of ways.
Understanding how to design consistent convex surrogates for different target losses and properties we wish to learn, and when this is impossible, can lead to understanding the tradeoffs that are often made in the current ad-hoc design of surrogates.
These ad-hoc surrogates are consistent with respect to \emph{some property}, although it may not be the one we wish to predict, either directly or indirectly.
As a general prediction task arises, one might be able to use property elicitation to understand what question they are actually answering (in the best case, with sufficient data that matches the real-world data distribution) by minimizing their surrogate loss and how it differs from the question they are actually trying to answer.
Of course, improvements in general machine learning algorithms can be put to many uses in many societal contexts, so a more specific analysis of potential broader impacts is very difficult to forecast.

@@

%\begin{ack}
\raf{ACKS: adam, nishant, nsf grants/fellowship}
%\end{ack}
